---
title: "HW5_Brown_Nick"
author: "Nick Brown"
date: "10/1/2019"
output:
  html_document:
    df_print: paged
---

Problem 1
--Knit to HTML

Problem 2
--Calculate sum of squares errors

```{r ss_errors}

set.seed(12345)
y <- seq(from=1, to=100, length.out = 1e8) + rnorm(1e8)

# Part a - iteration loop

system.time({
sequence <- length(y)
mean_y = mean(y)
sum_squares = 0
for(i in 1:sequence) {
  sum_squares = sum_squares + ((y[i] - mean_y)^2)
}
})
print(sum_squares)

# Part b - vector operations
system.time({sum((y - mean(y))^2)})
sum((y - mean(y))^2)
```

Problem 3
--Given X and ~h below, implement the above algorithm and compare the results with lm(h~0+X). State the
tolerance used and the step size, α.

```{r regression}

set.seed(1256)
theta <- as.matrix(c(1,2),nrow=2)
X <- cbind(1,rep(1:10,10))
h <- X%*%theta+rnorm(100,0,0.2)

#lm(h~theta + X)
```

Problem 4
--what are you to do?? Can you explain what is going on?

Beta-hat coefficient is typically calculated as (X'X)^-1 X'but when this expression is multiplied by X, then it creates the hat matrix. In this expression, however, the expression is multiplied by the y matrix instead.

Problem 5
--Compute the following: 
y = p + AB^−1 (q − r)

```{r need_for_speed}

#set.seed(12456)
#G <- matrix(sample(c(0,0.5,1),size=16000,replace=T),ncol=10)
#R <- cor(G) # R: 10 * 10 correlation matrix of G
#C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
#id <- sample(1:16000,size=932,replace=F)
#q <- sample(c(0,0.5,1),size=15068,replace=T) # vector of length 15068
#A <- C[id, -id] # matrix of dimension 932 * 15068
#B <- C[-id, -id] # matrix of dimension 15068 * 15068
#p <- runif(932,0,1)
#r <- runif(15068,0,1)
#C<-NULL #save some memory space

#object.size(A)
#object.size(B)

#system.time({y <- p+A*B**-1*(q-r)})

```

Part a:

The bytes size of A is 112347224 bytes
The bytes size of B is 1816357208 bytes

Part b.
How would you break apart this compute, i.e., what order of operations would make sense? We would first start with the B array and raise to the power of -1. (Unfortunately, my code would not compute this.) Then we would multiply this figure by A. Next we would subtract r from q, then multiply this times the aforementioned expression. Last, we would add p.

Are there any mathmatical simplifications you can make? The formula seems straightforward, therefore, it doesn't appear that a simplification can be made.

Is there anything about the vectors or matrices we might take advantage of? Yes, the result from q-r should be its own vector so that multiplying AB and (q-r) can occur use matrix algebra.

Part c.
