---
title: "HW8_Brown"
author: "Nick Brown"
date: "10/30/2019"
output: pdf_document
---

```{r setup, include=FALSE}
# Install tidyverse and dplyr
options(repos = c(CRAN="http://cran.us.r-project.org"))
options(tinytex.verbose = TRUE)
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("gridExtra")
library(tidyverse)
library(knitr)
library(psych)
library(ggplot2)
library(gridExtra)
```

##Problem 1:

Completed Swirl Exploratory Data Analysis lessons 1 through 10.

##Problem 2:

HW8 created as necessary.

##Problem 3:

Clean data from the World Bank

```{r worldbank}
zip_edu_url <- "https://databank.worldbank.org/data/download/Edstats_csv.zip"

############# Read all files from a .zip file ##############################
#(citation: https://hydroecology.net/downloading-extracting-and-reading-files-in-r/)

# create a temporary directory
td = tempdir()
# create the placeholder file
tf = tempfile(tmpdir=td, fileext=".zip")
# download into the placeholder file
download.file(zip_edu_url, tf)

# get the name of the first file in the zip archive
fname = unzip(tf, list=TRUE)$Name[1]
# unzip the file to the temporary directory
unzip(tf, files=fname, exdir=td, overwrite=TRUE)
# fpath is the full path to the extracted file
fpath = file.path(td, fname)

# stringsAsFactors=TRUE will screw up conversion to numeric!
edu = read.csv(fpath, header=TRUE, row.names=NULL, 
             stringsAsFactors=FALSE)
####################################################################################
```

How many data points were there in the complete dataset?

There are 886,930 observations of 70 variables in the complete dataset

```{r clean_edu}
###### Clean Dataset #######################

clean_edu <- edu[!is.na(edu$X1970),1:70]
```

How many data points were there in the cleaned dataset?

There are 72,288 observations of 70 variables in the cleaned dataset

Choosing 2 countries, create a summary table of indicators for comparison.
```{r summary}
###### Choose two countries #######################

Korea <- subset(clean_edu, Country.Code == "KOR" )
Turkey <- subset(clean_edu, Country.Code == "TUR" )
korea_indicator <- summary(Korea$Indicator.Name)
turkey_indicator <- summary(Turkey$Indicator.Name)
kable(cbind(korea_indicator,turkey_indicator), caption="Summary table of indicators for comparison")
```


## Problem 4

Using base plotting functions, recreate the scatter plot shown in class with histograms in the margins.  You do not have to make the plot the same, just have a scatter plot with marginal histograms.  Demonstrate the plot using suitable data from problem 2.

```{r scat}
###### Scatterplot with marginal histograms #######################

Korea <- subset(clean_edu, Country.Code == "KOR" )
Turkey <- subset(clean_edu, Country.Code == "TUR" )

x=na.omit(Korea$X1970)
y=na.omit(Korea$X2000)

x=x[-(length(y)+1):-length(x)]

x=rnorm(x)
y=rnorm(y)

scatter.hist(x,y)

```



## Problem 5

Recreate the plot in problem 3 using ggplot2 functions.  Note: there are many extension libraries for ggplot, you will probably find an extension to the ggplot2 functionality will do exactly what you want.

```{r ggscat}
###### Scatterplot with marginal histograms #######################

Korea <- subset(clean_edu, Country.Code == "KOR" )
Turkey <- subset(clean_edu, Country.Code == "TUR" )

x=na.omit(Korea$X1970)
y=na.omit(Korea$X2000)

x=x[-(length(y)+1):-length(x)]

x=rnorm(x)
y=rnorm(y)

hist_top <- ggplot()+geom_histogram(aes(x))
empty <- ggplot()+geom_point(aes(1,1), colour="white")+
         theme(axis.ticks=element_blank(), 
               panel.background=element_blank(), 
               axis.text.x=element_blank(), axis.text.y=element_blank(),           
               axis.title.x=element_blank(), axis.title.y=element_blank())

scatter <- ggplot()+geom_point(aes(x,y))
hist_right <- ggplot()+geom_histogram(aes(y))+coord_flip()

grid.arrange(hist_top, empty, scatter, hist_right, ncol=2, nrow=2, widths=c(4, 1), heights=c(1, 4))

```

